---
title: "MERS"
author: "Kate Sabey"
date: "May 16, 2018"
output: 
    html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


## MERS Background

Middle East Respiratory Syndrome Corona Virus (MERS-CoV) is a single-stranded virus related to SARS and other coronaviruses that causes severe respiratory illness. We will use a data set on the spread of MERS-CoV available at: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv

```{r packages, include=FALSE}
library(ggplot2)
library(plotly)
library(lubridate)
```


## Upload Data

After downloading the "cases.csv" file to your working directory, load the data into R Studio

```{r mers, include=TRUE}
mers <- read.csv("cases.csv")
```


## Manipulate Data

Format a new onset column to contain onset values as dates.

```{r data, include=TRUE}
mers$onset2 <- ymd(mers$onset)
```

Calculate the days elapsed between the date of onset and the start of the epidemic.

```{r data2, include=TRUE}
day0 <- min(na.omit(mers$onset2))
```

Create a column containing values for the time since the start of the epidemic.

```{r data3, include=TRUE}
mers$epi.day <- as.numeric(mers$onset2 - day0)
```


## Plot a MERS Epidemic Curve

```{r epi.day, include = TRUE}
epi.curve <- ggplot(data=mers) +
  geom_bar(mapping=aes(x=epi.day)) +
  labs(x='Epidemic day', y='Case count', title='Global count of MERS cases by date of symptom onset', 
       caption="Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv")

ggplotly(epi.curve)
```

